Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-30 paper.qmd
Change the projections for Five Gaussian clusters

## First projection
projection <- cbind(
    c(0.46477,-0.02024,0.56378,0.39736),
    c(0.15607,-0.54876,-0.44090,0.41505))

@@

axes_obj <- gen_axes(
  data = scaled_gau_data, 
  proj_data = projected_df, 
  limits = 0.3,
  axis_pos_x = -0.4,
  axis_pos_y = -0.1)

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.45, -0.2)) +
  ylim(c(-0.15, 0.18)) +
  interior_annotation("a1", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

axes_obj <- gen_axes(
  data = scaled_gau_data, 
  proj_data = projected_df, 
  limits = 0.3,
  axis_pos_x = 0.1,
  axis_pos_y = -0.1)

@@

five_gau_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(0.05, 0.25)) +
  ylim(c(-0.15, 0.25)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

## First projection
projection <- cbind(
  c(0.02743,0.21427,0.62893,-0.49816),
  c(-0.70709,0.42156,-0.02593,0.10965))

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  limits = 0.3,
  axis_pos_x = -0.2,
  axis_pos_y = -0.85)

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.25, 0.1)) +
  ylim(c(-0.9, -0.55)) +
  interior_annotation("b1", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  limits = 0.3,
  axis_pos_x = 0.08,
  axis_pos_y = -0.08)

@@

five_gau_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(0, 0.4)) +
  ylim(c(-0.15, 0.3)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

## First projection
projection <- cbind(
    c(0.21889,-0.25701,-0.19529,-0.73412),
    c(-0.63168,0.27992,-0.42895,-0.17223))

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  limits = 0.3,
  axis_pos_x = 0.1,
  axis_pos_y = 0.05)

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(0.05, 0.35)) +
  ylim(c(0, 0.35)) +
  interior_annotation("c1", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

five_gau_proj_pacmap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.8,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.75, -0.35)) +
  ylim(c(-0.75, -0.4))+
  interior_annotation("c2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

```{r}
#| label: fig-five-gau-nldr-layouts
#| fig-height: 5
#| fig-width: 15
#| fig-pos: H
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"

>>>>

## First projection
# projection <- cbind(
#     c(0.46477,-0.02024,0.56378,0.39736),
#     c(0.15607,-0.54876,-0.44090,0.41505))
projection <- cbind(
  c(-0.00215,-0.68905,-0.04778,-0.54223),
  c(0.42558,-0.23854,-0.63659,0.35753))

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  limits = 0.2,
  axis_pos_x = -0.12,
  axis_pos_y = -0.12)

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.41, -0.15)) +
  ylim(c(0.18, 0.52)) +
  interior_annotation("a1", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

@@

five_gau_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.75, -0.35)) +
  ylim(c(-0.75, -0.4)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

## First projection
# projection <- cbind(
#   c(0.02743,0.21427,0.62893,-0.49816),
#   c(-0.70709,0.42156,-0.02593,0.10965))
projection <- cbind(
  c(-0.00215,-0.68905,-0.04778,-0.54223),
  c(0.42558,-0.23854,-0.63659,0.35753))

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  limits = 0.2,
  axis_pos_x = 0.45,
  axis_pos_y = -0.28)

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.41, -0.15)) +
  ylim(c(0.18, 0.52)) +
  interior_annotation("b1", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

@@

five_gau_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.75, -0.35)) +
  ylim(c(-0.75, -0.4)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

## First projection
projection <- cbind(
  c(-0.00215,-0.68905,-0.04778,-0.54223),
  c(0.42558,-0.23854,-0.63659,0.35753))

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  limits = 0.2,
  axis_pos_x = -0.37,
  axis_pos_y = 0.22)

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.41, -0.15)) +
  ylim(c(0.18, 0.52)) +
  interior_annotation("c1", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

five_gau_proj_pacmap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.8,
    alpha = 0.15) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.75, -0.35)) +
  ylim(c(-0.75, -0.4))+
  interior_annotation("c2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

```{r}
#| label: fig-five-gau-nldr-layouts
#| fig-height: 10
#| fig-width: 15
#| fig-pos: H
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
